---
title: "DS311 - R Lab Assignment"
author: "Ana Paula Felix de Queiroz"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document:
    theme: united
    highlight: tango
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Assignment 1

* In this assignment, we are going to apply some of the build in data set in R for descriptive statistics analysis.
* To earn full grade in this assignment, students need to complete the coding tasks for each question to get the result.
* After finished all the questions, knit the document into HTML format for submission.

### Question 1

Using the **mtcars** data set in R, please answer the following questions.

```{r}
# Loading the data
data(mtcars)

# Head of the data set
head(mtcars)
```

a. Report the number of variables and observations in the data set.

```{r}
# Enter your code here!
dim(mtcars)

# Answer:
print("There are total of 11 variables and 32 observations in this data set.")
```



b. Print the summary statistics of the data set and report how many discrete and continuous variables are in the data set.

```{r}
# Enter your code here!

summary(mtcars)

continuous <- sum(sapply(mtcars, is.double))
discrete <- ncol(mtcars) - continuous
continuous
discrete
# Answer:
print("There are 0 discrete variables and 11 continuous variables in this data set.")
```



c. Calculate the mean, variance, and standard deviation for the variable **mpg** and assign them into variable names m, v, and s. Report the results in the print statement.

```{r}
# Enter your code here!
attach(mtcars)
m=mean(mpg)
v=sd(mpg)^2
s = sd(mpg)

print(paste("The average of Mile Per Gallon from this data set is ",m  , " with variance ",v  , " and standard deviation", s , "."))
```



d. Create two tables to summarize 1) average mpg for each cylinder class and 2) the standard deviation of mpg for each gear class.

```{r}
# Enter your code here!

library(dplyr)

# Create a table of average mpg for each cylinder class
mpg_by_cyl <- mtcars %>% 
  group_by(cyl) %>% 
  summarise(avg_mpg = mean(mpg))

mpg_by_cyl

# Create a table of standard deviation of mpg for each gear class
mpg_by_gear <- mtcars %>% 
  group_by(gear) %>% 
  summarise(sd_mpg = sd(mpg))

# Print the table
print(mpg_by_gear)





```



e. Create a crosstab that shows the number of observations belong to each cylinder and gear class combinations. The table should show how many observations given the car has 4 cylinders with 3 gears, 4 cylinders with 4 gears, etc. Report which combination is recorded in this data set and how many observations for this type of car.

```{r}
# Enter your code here!
crosstab = table(mtcars$cyl,mtcars$gear,
                 dnn = c("Cyl","gears"))
crosstab

print("The most common car type in this data set is car with 8 cylinders and 3 gears. There are total of 12 cars belong to this specification in the data set.")
```


*** 

### Question 2

Use different visualization tools to summarize the data sets in this question.

a. Using the **PlantGrowth** data set, visualize and compare the weight of the plant in the three separated group. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your findings.

```{r}
# Load the data set
data("PlantGrowth")

# Head of the data set
head(PlantGrowth)

# Enter your code here!

library(ggplot2)
ggplot(PlantGrowth, aes(x=group, y=weight, fill=group)) + 
  geom_boxplot() + labs(title = "Weight of Plants by Group", x = "Group", y = "Weight")
```

Result:

=> Report a paragraph to summarize your findings from the plot!

The box plot shows that weights of plants for group TRT2 is higher than the other two with a median at 5.4. While the group Trt1 has the lower median around 4.6. Group CTRL seems to have the largest spread of the 50% of the data. 
The distribution of group ctrl seem to be skewed to the write while group trt2 seem to be skewed to the left.We are also able to identify that trt1 has outliers. 

b. Using the **mtcars** data set, plot the histogram for the column **mpg** with 10 breaks. Give labels to the title, x-axis, and y-axis on the graph. Report the most observed mpg class from the data set.

```{r}
hist(mtcars$mpg, breaks = 10,
     main = "MPG distribution",
     xlab = "MPG",
     ylab = "Frequency")


print("Most of the cars in this data set are in the class of 10 mile per gallon.")
```



c. Using the **USArrests** data set, create a pairs plot to display the correlations between the variables in the data set. Plot the scatter plot with **Murder** and **Assault**. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your results from both plots.

```{r}
# Load the data set
data("USArrests")

# Head of the data set
head(USArrests)

# Enter your code here!


pairs(USArrests, main = "Pairs Plot")

plot(USArrests$Murder, USArrests$Assault,
     main = "Scatter Plot of Murder and Assault",
     xlab = "Murder", ylab = "Assault")
```

Result:

=> Report a paragraph to summarize your findings from the plot!
It seems like murder and Assault have the highest positive correlation. 
We could say that there is a moderate positive correlation between Rape and murder as well.
It seems like Urban Population is the variable with the lowest correlation with Assault and murder. 



***

### Question 3

Download the housing data set from www.jaredlander.com and find out what explains the housing prices in New York City.

Note: Check your working directory to make sure that you can download the data into the data folder. 

```{r, echo=FALSE}
dir.create("data")
# Load and clean the housing data set 
download.file(url='https://www.jaredlander.com/data/housing.csv',
              destfile='data/housing.csv', mode='wb')
housingData <- read.csv('data/housing.csv')
housingData <- subset(housingData, 
                      select = c("Neighborhood", "Market.Value.per.SqFt", "Boro", "Year.Built"))
housingData <- na.omit(housingData)                   
```

a. Create your own descriptive statistics and aggregation tables to summarize the data set and find any meaningful results between different variables in the data set.

```{r}
# Head of the cleaned data set
head(housingData)
unique(housingData$Boro)
summary(housingData)


# Enter your code here!
#table discribing mean and median grouped by value by var Boro
library(dplyr)
housingData %>% 
  group_by(Boro) %>% 
  summarize(mean_value = mean(Market.Value.per.SqFt), 
            median_value = median(Market.Value.per.SqFt))


housingData %>%
  group_by(Neighborhood) %>%
  summarize(mean_value = mean(Market.Value.per.SqFt),
            median_value = median(Market.Value.per.SqFt)) %>%
  arrange(desc(mean_value))

summary(housingData$Market.Value.per.SqFt)


```

b. Create multiple plots to demonstrates the correlations between different variables. Remember to label all axes and give title to each graph.

```{r}
# Enter your code here!
library(ggplot2)
ggplot(housingData, aes(x = Boro, y = Market.Value.per.SqFt)) +
  geom_point() +
  labs(title = "Market Value by Boro",
       x = "Boro",
       y = "Market Value per SqFt")
top10_neighborhoods <- housingData %>%
  group_by(Neighborhood) %>%
  summarize(mean_value = mean(Market.Value.per.SqFt)) %>%
  arrange(desc(mean_value)) %>%
  top_n(5, mean_value)


ggplot(top10_neighborhoods, aes(x = reorder(Neighborhood, mean_value), y = mean_value)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Top 5 Neighborhoods by Market Value per SqFt",
       x = "Neighborhood", y = "Mean Market Value per SqFt")


ggplot(housingData, aes(x = Year.Built, y = Boro)) +
  geom_boxplot() +
  labs(title = "Distribution of Boro by Year Built",
       x = "Year Built",
       y = "Boro")

```

c. Write a summary about your findings from this exercise.

If we look at the variables Boro and market value we can see that Mahattan has the highest market value per square ft, queens and brooklyn seem to have a similar Market price exept for outliers. its followed by Bronx and then Staten Island. 

Among all the Neighboors Midtown, Flatiron, Midtown West , uper East Side and Chelsea hank on the top highest mean for Market value per square ft

We see that broklyn has one observation as the oldest building, and manhatan has the largest distribution of buildings age with the median around 1960.





